---
title: "DS6372_Project1_LifeExpectancy"
author: 'Bala Dakshinamoorthi, Daanesh Ibrahim and Jason Rupp'
date: "6/14/2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
library(pander)
library(tidyverse)
library(naniar)
library(gridExtra)
library(corrplot)
library(dplyr)
library(ggplot2)
library(GGally)
library(plotly)
library(fastDummies)
library(car)
library(class)
library(caret)
library(MASS)
library(tree)

# Read CSV

life_df <- read_csv("data/LifeExpectancyData.csv")

# life_df <- read_csv(file.choose())

```


## Introduction

In the global health community, a common measure of the health status of a nation is the life expectancy of a person at birth. This measure can tell a lot about the strength of the nation's health care system and there are great many contributing factors which go into determining this measure. The difficulty with determining life expectancy is that if it is a measure of a population that is living, it will always an estimate. The actual life span or expected life span of a population will take a number of years to determine, as the group would have to expire. To further complicate matters, these estimates are based on whether the conditions of a health system or the relative health status of a nation will remain constant. Due to these factors, we must rely on statistical methods to derive these measures, as they can be very useful to the global health community. The purpose of this project is to use a set of data with roughly 20 features related to the health care systems of 193 nations to develop statistical models which may best predict the life expectancy of a nation, and satisfy a subset of other objectives, that will be elaborated on later. We will begin by presenting the data then briefly explain steps used to evaluate said data for model selection.



## Data Description 

The data set used for this project was obtained from Kaggle, the link for which can be found in the appendix accompanied with a full description of each feature as described on the website, and the statistical summary of the data for reference. The overall data set contained 22 columns, 1 column that identified the country, 1 which held the life expectancy of the country and 20 features related to the health and economic status of 193 nations from a 15 year period, between 2000-2015. The user that posted the data claims to have obtained the health related features from the Global Health Observatory (GHO) data repository under World Health Organization (WHO) and the economic features from the United Nations website. Additionally, this user did note that it appeared to be free of errors, though there were missing values in some of the features. The missing values will addressed later in the discussion. Most of the attributes are self explanatory with regard to the value and the unit, here are some key points about the data the will bring context into interpretation of the models further in the paper. Polio, Hepatitis B, Diphtheria (Diphtheria tetanus toxoid and pertussis - DTP3) refer to the immunization coverage among 1-year-olds (%). The infant, under-five, HIV/AIDS features refer to the number of deaths per 1000 population; HIV/AIDS in the age range of 0-4 years. Adult Mortality is the probability  of both sexes of dying between 15 and 60 years per 1000 population. Measles is the number of reported cases per 1000 population. One thing of note that was noticed about the data set description was that the "thinness 1-19 years" feature said in the description that it was the	prevalence of thinness among children and adolescents for Age **10** to 19 (%), so this is a bit of an ambiguous attribute which could refer to either age, either way, it was noted an analysis continued. 

```{r}
# Make initial df with only 2014

life_df <- 
  life_df %>%
  filter(Year == 2014)

# Fix column header

for (i in 1:length(names(life_df))) {
  ifelse(
    grep(" ",names(life_df)[i]) == TRUE,
    names(life_df)[i] <- gsub(" ", "_", names(life_df)[i]),
    next
  )
}

# Drop all rows with NAs

life <- life_df %>% 
  drop_na()

# Create new data frame with factors on immunizations

life_new <- life[,c(-1,-2)] %>% 
  mutate(Hepatitis_B = ifelse(Hepatitis_B < 86, "<86% Immunized", ">=86% Immunized"),
         Polio = ifelse(Polio < 86, "<86% Immunized", ">=86% Immunized"),
         Diphtheria = ifelse(Diphtheria < 86, "<86% Immunized", ">=86% Immunized"),
         Hepatitis_B = as.factor(Hepatitis_B),
         Polio = as.factor(Polio),
         Diphtheria = as.factor(Diphtheria))

```

## Exploratory Data Analysis

Before statistical modeling could begin, a thorough investigation of the data was necessary. It is important to understand the nature of the data, before applying modeling techniques. As previously stated, first we set out to identify what values were missing in the data set, and come to a decision on how to proceed. 

```{r echo=FALSE}
# Show quick plot of NAs in each column

gg_miss_var(life_df)
```

As the graphic above shows, there are several missing values as pointed out in the website. In order to solve this issue, a multi-prong approach was utilized. One approach was simply to drop all the missing values from the model. This was justified with the reasoning that the modeling will be done on only the data from the year 2014, this act alone cut nearly 80-90% from the entire data set, taking almost 2,800 of the data points out of the set from the beginning, it seemed inconsequential to eliminate a few more rows, 52 more rows dropped in total. Another approach was to reference some of the easy values to fill in, mainly the population and GDP portions of the data. These missing values (GDP and population) were easily found from reputable sources, namely from the World Bank website. The additional missing values were filled in using the median of the attribute from where they came. This approach certainly isn't perfect, and could easily yield absurd values with regard to the country itself, but it allowed the utilization of more degrees of freedom for some of the models that were designed. When final modeling occurred, the dataset that was given was used, and the NA values were dropped.

Before we proceed with the preliminary findings from data analysis, there is fact that should be noted. There is a stark difference in mean life expectancy between the developed and developing nations. Our data set supported this notion, as one can see below, the mean life expectancy for each type of country is noted, along with the plot to show the range.

```{r echo=FALSE}
pander(aggregate(life_new$Life_expectancy, by=list(Status=life_new$Status), FUN=mean))
```

```{r echo=FALSE}
life_plot <-  ggplot(life_new, aes(x=Status, y = Life_expectancy, fill = Status)) +
  geom_boxplot() +
  scale_fill_manual(values=c("forestgreen", "darkorange")) +
  labs(x = "Status", y = "Life Expectancy (Age)") +
  theme(legend.position = "none")
ggplotly(life_plot)
```

To bolster the argument, a one-way anova was performed with only status as an explanatory, and though it is not in line with the scope of this research, the result showed that the difference observed in this dataset is statistically relevant, p-value<0.001. This could complicate further analysis, but we will proceed with analysis, and will revisit this concept later in the discussion.

To investigate the relationships between each of the predictor variables, and our desired response variable. A correlation plot of the numerical data was created which guiding much of the initial stages of data analysis.

```{r Correlation Numeric, echo=FALSE}
life_numerical <- life_new %>% 
  select_if(is.numeric)

ggcorr(life_numerical, 
       label = T, 
       label_size = 2,
       label_round = 2,
       hjust = 1,
       size = 3, 
       color = "black",
       layout.exp = 5,
       low = "forestgreen", 
       mid = "gray95", 
       high = "darkorange",
       name = "Correlation")
```

With this plot a number of quick observations were possible, and it was the goal to identify relationships between variables. We were able gain some inference on our data set as it relates to the response variable. The above plot shows that life expectancy has a strong positive correlation with Schooling and Income_composition_of_resources. Given this information, we were able to develop a rudimentary model involving these variables. Moving on, we also found that life expectancy has a strong negative correlation with adult mortality which is understandable and intuitive. The description of the data notes that this feature is the probability of death between the ages of 15-60, and since if mortality rate in adults is high, life expectancy would be lower, this finding supports what we would expect. 

There are a few points to note that was discovered about the immunization statistics particularly Hepatitis B, Polio and Diphtheria. There was a very large range in the amount of coverage for each of the countries. Many countries have values which lie closer to 100, but some are in the single digits. According to the World Health Organization in 2018, 86% of children in the world are receiving immunizations protecting them from these diseases. For some modeling, 86% was used as a benchmark to turn these columns into factors.

```{r Hepatitis_B Variable, echo=FALSE}
life_plot2 <-  ggplot(life_new, aes(x=Hepatitis_B, y = Life_expectancy, fill = Hepatitis_B)) +
  geom_boxplot() +
  scale_fill_manual(values=c("forestgreen", "darkorange")) +
  labs(x = "Hepatitis B Immunizations", y = "Life Expectancy (Age)") +
  theme(legend.position = "none")
ggplotly(life_plot2)
```

The above graphic shows that more than 1/3 of countries have less than 86% immunized for Hepatitis B. The Life Expectancy of the countries with greater than or equal to 86% Immunized is higher than the countries which have less that 86% Immunized. Note that the median (2nd Quartile) for >=86% immunized is skewed upwards towards the 3rd quartile.

```{r Polio Variable, echo=FALSE}
life_plot3 <-  ggplot(life_new, aes(x=Polio, y = Life_expectancy, fill = Polio)) +
  geom_boxplot() +
  scale_fill_manual(values=c("forestgreen", "darkorange")) +
  labs(x = "Polio Immunizations", y = "Life Expectancy (Age)") +
  theme(legend.position = "none")
ggplotly(life_plot3)
```

For Polio Immunizations, higher life expectancy resides with the countries that have >=86% Immunized. Note that the median (2nd Quartile) for >=86% immunized is skewed upwards towards the 3rd quartile.

```{r Diptheria Variable, echo=FALSE}
life_plot4 <-  ggplot(life_new, aes(x=Diphtheria, y = Life_expectancy, fill = Diphtheria)) +
  geom_boxplot() +
  scale_fill_manual(values=c("forestgreen", "darkorange")) +
  labs(x = "Diphtheria Immunizations", y = "Life Expectancy (Age)") +
  theme(legend.position = "none")
ggplotly(life_plot4)
```

For Diphtheria Immunizations, higher life expectancy resides with the countries that have >=86% Immunized. Note that the median (2nd Quartile) for >=86% immunized is skewed upwards towards the 3rd quartile. 

These immunization findings really give an overview in the difference in health systems. Further investigations noted that all of the developed countries have >=86% of their people immunized for Polio, and Diphtheria, and A larger percentage of Developed countries have >=86% of their people immunized for Hepatitis_B. Most other findings during preliminary data analysis was unremarkable. The next section will present the objectives of this analysis in depth and outline the methods used to solve these goals.

## Objective 1:

```{r Splitting}
# setting seed ,splitting data set into train test sets, 70/30 further later model assessment
set.seed(1234)

life_new_idx <- 
  sample(
    1:dim(life_new)[1],
    round(dim(life_new)[1] * 0.7),
    replace=F
    )

life_new_train <- life_new[life_new_idx,]

life_new_test <- life_new[-life_new_idx,]
```


#### Primary Model

The first objective was to build a model with the main goal to identify key relationships which is highly interpretable. To create a simple model, we first evaluated the data set for multicolinearity to remove any redundancy. The code for this can be found in the appendix, accompanied with a full explanation of the reasoning for removing the variables, however Infant_deaths, Income_composition_of_resources, thinness_5-9_years, and percentage_expenditure were removed. With colinearity addressed, we added the remaining of variables to the model, in hope of identifying anything of note with regard to the coefficients or the p-values. Below is a summary of the model analysis.

```{r Model Creation}
life_new_vif<- life_new[-c(6,10,18:19)]
life_lm <- lm(formula = Life_expectancy ~., data = life_new_vif)

```

From the model we found the negative coefficients are Status - Developing, Adult_Mortality, infant_deaths, Hepatitis_B>=86% Immunized, Polio>=86% Immunized, HIV/AIDS, and thinness__1-19_years. This is important because it tells us that we could typically expect to see an decrease in life expectancy as these values rise. Diphtheria>=86% Immunized is expected to raise Life_expectancy while Hepatitis_B>=86% Immunized and Polio>=86% Immunized are associated with a decrease in life expectancy, surprisingly. It is our opinion that we believe from this model, Adult_Mortality, HIV/AIDS, and Schooling are seen as the best predictors of life expectancy to create a highly interpretable model. 

According to the Adj. R-Squared from the entire model, much of the variation in life expectancy can be explained by the explanatory values of the model. This is a very starting point, though we wanted to perform various feature selection techniques to see if there was an improvement. We applied forward, backward and stepwise feature selection techniques to the data to gain insights, see the appendix for code, the output is very long.


According to each feature selection, the 3 variables that have the most significance with regards to Life_expectancy are Adult_Mortality, HIV/AIDS, and Schooling. They have the lowest p-values in each of the selection methods and some were chosen by all model selection methods. With interpretation in mind, these variables were used to create a simple model, from a 70/30 split of the data with a training and test, for comparison with other models.

```{r High p-value variables only}
life_highp <- lm(formula = Life_expectancy ~ Adult_Mortality + `HIV/AIDS` + Schooling, data = life_new_train)

summary(life_highp)
```

```{r}
predictionz <- life_highp %>% predict(life_new_test)

# Compute the prediction error ASE
life_highp_ase <- RMSE(predictionz, life_new_test$Life_expectancy)^2

life_pred <- predict(life_highp, life_new_test, interval="confidence")

life_highp_means <- summary(life_pred)[4,]
```


```{r echo=FALSE}
# Show residual plots
par(mfrow=c(2,3))
plot(life_highp, which = c(1:5))
```

The simple model residual plot show there is relative normality, additionally there doesn't seem to be any points to worry about with regard to Cook's distance. There are some high leverage points, however the residuals are low and we are not concerned. Below is the interpretation of the model.

|Coefficients:  |          |
|:--------------|:---------|
|Adult_Mortality| -0.031886|
|HIV/AIDS       | -0.854541|
|Schooling      |  1.504618|

```{r echo=FALSE}
pander(life_highp_means)
```

We are 95% confident that the interval [68.59, 72.01] will contain mean Life_expectancy for the model.

Interpretation: 

Adult_Mortality - For every one unit increase of adult mortality, the Life_expectancy is expected to decrease by -0.031, holding all the other variables constant. This predictor is a rather obvious one to include in terms of practical and statistical significance. It was included in the model specifically because it is significant statistically with regard to the data. The practical significance of this predictor may be questionable, as we can only go by the kaggle description, it says it is the probability of dying between 15-60 years per 1000 people. It is not very clear where this probability is calculated from, and we may not necessarily want to rely on a calculated or predicted value in our model but we have included it.

HIV/AIDS - For every one unit increase of HIV/AIDS (Deaths per 1000 live births HIV/AIDS 0-4 years), the life expectancy is expected to decrease by -0.85 holding all the other variables constant. This predictor is a better with regard to practical significance, as it can be empirically measured and it could also garner further study as to how to reduce this number in society. This variable could have a number of confounding effects, so it may be difficult really assess how practical significance.

Schooling - For every additional year of schooling the life expectancy is expected to increase by 1.50 holding all the other independent variables constant. Practical significance of this variable is evident; it would be relatively easy to measure with regard to a population, and by this model, will likely have an impact on the life expectancy of a country. The confounding effect of this variable is that schooling in nations is involved with the public infrastructure in place in the nation, and hospitals would be a part of this infrastructure. So a higher amount of schooling would likely mean that a nation has better funding to do so, and would also more than likely, have the ability to invest in health care. 


#### Second Model

```{r include=FALSE}
# Removing country name, and the year, keeping all other variables.

life_sw <- life[,c(-1:-2)]

# splitting new df on same indexes as selected before for head to head comparison

life_sw_train <- life_sw[life_new_idx,]

life_sw_test <- life_sw[-life_new_idx,]
```

The next objective of our study was to construct a model without regard to interpretability. The second model did not treat the immunizations for Hep B, Polio, and TDAP as factors during feature selection, but left them as the percentages they represented. The goal was to use all of the variables to perform stepwise feature selection to identify attributes, then assess what was selected.


```{r}
set.seed(123)

# Training control for feature selection

train.control <- trainControl(method = "cv", number = 10)

# Stepwise regression feature selection

step.model <- train(Life_expectancy ~., data = life_sw_train,
                    method = "lmStepAIC", 
                    trControl = train.control,
                    trace = FALSE
                    )

# Final model coefficients
step1_cof <- step.model$finalModel[1]

# Adult_Mortality, infant_deaths, percentage_expenditure, under-five_deaths, Income_composition_of_resources left, will check vifs now
```

The initial stepwise selection was complete, the following attributes were left, Adult_Mortality, infant_deaths, percentage_expenditure, under-five_deaths, Income_composition_of_resources, we now checked the VIF for the variables exhibiting multicolinearity. Initial check of the VIFs show high colinearity between infant_deaths and under-five_deaths as we saw before, under-five deaths was again dropped for this model.

```{r}

step1_lm <- lm(formula = Life_expectancy ~ Adult_Mortality + infant_deaths + percentage_expenditure + 
    `under-five_deaths` + Income_composition_of_resources, 
    data = life_sw_train)

# check vifs

step1_vif <- vif(step1_lm)

pander(step1_vif)
```


```{r}

final_step1 <- lm(formula = Life_expectancy ~ Adult_Mortality + infant_deaths + percentage_expenditure + 
     Income_composition_of_resources, 
    data = life_sw_train)
# check vifs

final_step1_vif <- vif(final_step1)

pander(final_step1_vif)

```

The variables that were included in the final model were adult mortality, infant deaths, percentage expenditure, income composition of resources. Now we performed the assessment of residual diagnostics with regard to this model.

```{r echo=FALSE}
par(mfrow=c(2,3))
plot(final_step1, which = c(1:5))
```

Initial evaluation of this model showed a high leverage, outlier point. Upon further investigation, it is shown that the row has a value of 957 for infant deaths, this seems unusually high, and grounds to drop the row and re-run the model. In addition further research revealed what country this value was pointing to, which was India. A quick Google search had the value of around 36, so this variable is questionable. Even if this is legitimate, which seems unlikely, it is far enough of an outlier to be grounds to drop.

```{r}
final_step <- lm(formula = Life_expectancy ~ Adult_Mortality + infant_deaths + percentage_expenditure + 
     Income_composition_of_resources, 
    data = life_sw_train[-65,])

final_step_sum <- summary(final_step)

par(mfrow=c(2,3))
plot(final_step, which = c(1:5))
```

```{r}
predictionz2 <- final_step %>% predict(life_sw_test)

# Compute the prediction error ASE
final_step_ase <- RMSE(predictionz2, life_sw_test$Life_expectancy)^2
```

Our second model shared only one attribute with our highly interpretable model which was adult mortality. Additionally, the second model included income composition of resources, which we saw during EDA to be most highly correlated with life expectancy. The reason this model loses interpretability it seems is due to the nature of the variables. Percentage expenditure was a value that was likely calculated by the user that uploaded the data, and the nature of the income composition of resources seems unreliable. This is a value that is supposed to represent a rather obscure measure, and could confound the result, however we saw it had the highest correlation with life expectancy.

In spite of these factors, our primary model, which we built by a combination of feature selection and intuition performed better on the reserved test set than the second model using stepwise feature selection only. The primary model had an average squared error of `r round(life_highp_ase,2)` and the second model was `r round(final_step_ase,2)`, which isn't a wide margin, but an interesting finding. 

## Objective 2

The last objective of this discussion was to choose a nonparametric modeling strategy to make a prediction on this data set. The method of choice here was k-nearest neighbors regression. This method is relatively easy to understand, with regard to k-nn for classification, and very similar methodology is used in regression, it is also easy to create a model with software. The goal is to use points nearest where the prediction is being made to guide the decision, the k, stands for how many neighbors will be used. To avoid overfitting, the features used for k-nn were what was selected from our second model, using stepwise selection. These features were used to model, and a range of k values were searched up to 10, to find the ideal number.

```{r}
# Fit the model on the training set
set.seed(1234)
knn_model <- train(
  Life_expectancy~., data = life_sw_train[,c(2,3,4,6,19)], method = "knn",
  trControl = trainControl("cv", number = 10),
  preProcess = c("center","scale"),
  tuneLength = 10
  )

# Make predictions on the test data
knn_predictions <- knn_model %>% predict(life_sw_test)

# Compute the prediction error ASE
knn_ase <- RMSE(knn_predictions, life_sw_test$Life_expectancy)^2
```

Surprisingly, the nonparametric k-nn regression performed better than either of the earlier regressions,`r round(life_highp_ase,2)` versus `r round(knn_ase,2)` for the k-nn model. 

	
## Final Conclusions

In terms of prediction, and with all of the models presented above, it would likely be a wise choice to use the k-nn regression, using 5 neighbors, given it yielded the lowest average squared error on the test set. It doesn't matter if the model can be interpreted if it can more accurately predict.

This was a very interesting study in which we were able to apply statistical methodologies to global socioeconomic factors to assess the effect on the response variable. From reading websites that pertain to global health, there always seem to be a myraid of factors which go into achieving and sustaining a high level of community health in the nation. Though the data set used for modeling was very small with regard to what is likely used on production level algorithms that these large organizations utilize, the basic strategdies are likely in place to deduce factors such as life expectancy. 

If there had been more time to complete this model, there are a few things we could have done differently. One thing that would be of interest pertained to the dataset. The website from which it came is known for practice, and it calls into question the legitimacy of the data set. It would be interesting to build different models with data sourced from a more reputable source with regard to global health data, which would undoubtedly be far more time intensive. Going to a larger data source may also open up unseen possibilties, such as acquiring more features for selection. 

Another interesting look at this dataset would be to investigate how life expectancy has changed for countries over time. This data has 15 years worth, so if time permitted, a full exploration could be had to see this change. What are the annual, or cyclical changes we can find, and how do all of these attributes flucuate over time, and how does that impact the life expectancy.








## Appendix 
 
Link to the dataset:
https://www.kaggle.com/kumarajarshi/life-expectancy-who
 
|---------------------------------------------STRAIGHT FROM KAGGLE----------------------------------------------------------------|

About this file
The Global Health Observatory (GHO) data repository under World Health Organization (WHO) keeps track of the health status as well as many other related factors for all countries The datasets are made available to public for the purpose of health data analysis. The dataset related to life expectancy, health factors for 193 countries has been collected from the same WHO data repository website and its corresponding economic data was collected from United Nation website. Among all categories of health-related factors only those critical factors were chosen which are more representative. It has been observed that in the past 15 years , there has been a huge development in health sector resulting in improvement of human mortality rates especially in the developing nations in comparison to the past 30 years. Therefore, in this project we have considered data from year 2000-2015 for 193 countries for further analysis. The individual data files have been merged together into a single dataset. On initial visual inspection of the data showed some missing values. As the datasets were from WHO, we found no evident errors. Missing data was handled in R software by using Missmap command. The result indicated that most of the missing data was for population, Hepatitis B and GDP. The missing data were from less known countries like Vanuatu, Tonga, Togo,Cabo Verde etc. Finding all data for these countries was difficult and hence, it was decided that we exclude these countries from the final model dataset. The final merged file(final dataset) consists of 22 Columns and 2938 rows which meant 20 predicting variables. All predicting variables was then divided into several broad categories:Immunization related factors, Mortality factors, Economical factors and Social factors.

|---------------------------------------------------------------------------------------------------------------------------------|

The following descriptions attached to these features are directly from the kaggle website.

Table: Full Data Description


|                   Feature                 |      Description                                                                                       |
|:------------------------------------------|:-------------------------------------------------------------------------------------------------------|
| Country	                                  |Country                                                                                                 |
| Year	                                    |Year                                                                                                    |
| Status	                                  |Developed or Developing status                                                                          |
| Life expectancy                           |Life Expectancy in age                                                                                  |
| Adult Mortality                           |Adult Mortality Rates of both sexes (probability of dying between 15 and 60 years per 1000 population)  |
| infant deaths  	                          |Number of Infant Deaths per 1000 population                                                             | 
| Alcohol		                                |Alcohol, recorded per capita (15+) consumption (in litres of pure alcohol)                              | 
| percentage expenditure	                  |Expenditure on health as a percentage of Gross Domestic Product per capita(%)                           |
| Hepatitis B                               |Hepatitis B (HepB) immunization coverage among 1-year-olds (%)                                          |
| Measles 	                                |Measles - number of reported cases per 1000 population                                                  |  
| BMI 	                                    |Average Body Mass Index of entire population                                                            |  
| under-five deaths 	                      |Number of under-five deaths per 1000 population                                                         |  
| Polio	                                    |Polio (Pol3) immunization coverage among 1-year-olds (%)                                                |  
| Total expenditure	                        |General government expenditure on health as a percentage of total government expenditure (%)            |    
| Diphtheria 	                              |Diphtheria tetanus toxoid and pertussis (DTP3) immunization coverage among 1-year-olds (%)              |  
| HIV/AIDS	                                |Deaths per 1 000 live births HIV/AIDS (0-4 years)                                                       |   
| GDP	                                      |Gross Domestic Product per capita (in USD)                                                              |
| Population	                              |Population of the country                                                                               |   
| thinness 1-19 years	                      |Prevalence of thinness among children and adolescents for Age 10 to 19 (%)                             |  
| thinness 5-9 years	                      |Prevalence of thinness among children for Age 5 to 9(%)                                                 |      
| Income composition of resources	          |Human Development Index in terms of income composition of resources (index ranging from 0 to 1)         | 
| Schooling                                 |Number of years of Schooling(years)                                                                     |    


```{r}
summary(life_df)
```


WHO 86% Immunized
Source: https://www.chop.edu/centers-programs/vaccine-education-center/global-immunization/diseases-and-vaccines-world-view
  
  
Primary Model VIF evaluation:
```{r check VIFs and compare to correlation plot to determine which columns to drop}
life_df_fctr2 <- life_new  #removing the country name, indicator vars
life_fctr2model <- lm(Life_expectancy~.,data=life_df_fctr2)  # . means all variable not mpg
vif(life_fctr2model)

```
Life Expectancy has a strong positive correlation with Schooling 
and Income_composition_of_resources but Schooling and Income_composition_of_resources have
a high correlation with one another. Income_composition_of_resources has a VIF of 11.35 (vs 7.47 for Schooling) so let's drop that.

Infant_deaths and under.five_deaths have a correlation of 1 (100%) which would
tell us that there is multicollinearity between them. Both have extremely high VIFs, so we should drop one of the variables. and I believe under 5 deaths should be dropped since it covers more ages and includes infant deaths within it.

Percentage_expenditure vs GDP is also highly correlated with one another with both values having high VIFs. GDP is a more well known datapoint and has a higher correlation with Life_expectancy, so let's remove percentage_expenditure.

thinness_5-9_years has a strong correlation with thinness_1-19_years and both have high VIFs. Let's drop thinness_5-9_years since thinness_1-19_years should have years 5-9 within it as well. 
  
  
```{r drop columns and rerun VIFs}

#removes Infant_deaths, Income_composition_of_resources, thinness_5-9_years, and percentage_expenditure.
life_new<- life_new[-c(6,10,18:19)]

life_df_fctr2 <- life_new 
life_fctr2model <- lm(Life_expectancy~.,data=life_df_fctr2)  
vif(life_fctr2model)

```


Primary Modeling Feature selection:
```{r Stepwise}
life_step <- lm(formula = Life_expectancy ~., data = life_new_vif)
life_step0 <- lm(formula = Life_expectancy ~1, data = life_new_vif)
```

```{r backwards}
life_backward <- step(life_step, direction = "backward")
summary(life_backward)
```

```{r forwards}
life_forward <- step(life_step0, scope = list(lower = life_step0, upper = life_step) ,direction = "forward")

summary(life_forward)
```

```{r both, include=FALSE}
life_both <- step(life_step, scope = list(lower = life_step0, upper = life_step) ,direction = "both")

summary(life_both)
```




