---
title: "Project 1"
author: "Daanesh"
date: "6/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown



```{r libs, include=FALSE}
# Required Libraries
library(tidyverse)
library(naniar)
library(gridExtra)
library(corrplot)
library(dplyr)
library(ggplot2)
library(GGally)
library(plotly)
library(fastDummies)
library(car)
```



```{r dataload}
# Read csv from project dir

# life_df <- read_csv("C:/Users/e005108/Downloads/datasets_12603_17232_Life Expectancy Data.csv")

# life_df <- read_csv("data/LifeExpectancyData.csv")  

life_df <- read_csv("data/LifeExpectancyData_w_regions.csv")

life_df <- 
  life_df %>%
  filter(Year == 2014)
head(life_df)
```


```{r}

life_df$Status <- as.factor(life_df$Status)

life_df$WB_region <- as.factor(life_df$WB_region)

# life_df_fctr <- dummy_cols(life_df, select_columns = "Status")

life_df_fctr <- dummy_cols(life_df, select_columns = c("Status", "WB_region"))


# This will drop the year, Status(character), Status(Developed/indicator) columns

# life_drop_cols <- c(2,3,23)

# life_df_fctr <- life_df_fctr[,-life_drop_cols]

life_df_fctr 

```



We need to convert "Status" to a factor then to an indicator variable

Status is binomial, so we only need one indicator variable, excess columns are dropped

We have the region for each country retrieved from the world bank website depicting the WHO regions around the world

We can added this and create indicator variables for the regions as well


```{r Header Change}
#Removing header spaces from the data frames

for (i in 1:length(names(life_df))) {
  ifelse(
    grep(" ",names(life_df)[i]) == TRUE,
    names(life_df)[i] <- gsub(" ", "_", names(life_df)[i]),
    next
  )
}


for (i in 1:length(names(life_df_fctr))) {
  ifelse(
    grep(" ",names(life_df_fctr)[i]) == TRUE,
    names(life_df_fctr)[i] <- gsub(" ", "_", names(life_df_fctr)[i]),
    next
  )
}

names(life_df)

names(life_df_fctr)

summary(life_df)
```







```{r Locate NAs}
gg_miss_var(life_df)

life_df %>%
  filter(
    Status == "Developed"
  ) %>%
  ggplot() +
  geom_col(
    aes(
      x = reorder(Country, Life_expectancy),
      y = Life_expectancy,
      fill = Status
    )
  ) +
  coord_flip() +
  theme(
    axis.text.y = element_text(size=6),
  )


life_df %>%
  filter(
    Status == "Developing"
  ) %>%
  ggplot() +
  geom_col(
    aes(
      x = reorder(Country, Life_expectancy),
      y = Life_expectancy,
      fill = Status
    )
  ) +
  coord_flip() +
  theme(
    axis.text.y = element_text(size=2),
  )


```

```{r Remove NAs and see the range of values for each column}

life <- life_df %>% 
  drop_na()

summary(life)

range(life$Life_expectancy)
```

22 Variables, 20 of them are Numerical, and 2 of them are Categorical.
Variables we should drop: Country, Year
Hepatitis.B has the Min Value and 1st Quartile difference of 76 which is too high (Factor?)
Polio has the Min value and 1st QUartile difference of 70 which is too high (Factor?)
Diphtheria has the min value and 1st quartile difference of 78 which is too high (Factor?)

According to the World Health Organization in 2018 they said that
86% of children in the world are receiving immunizations protecting them from these diseases.
Source: https://www.chop.edu/centers-programs/vaccine-education-center/global-immunization/diseases-and-vaccines-world-view
Let us use 86% as the benchmark to turn these columns into factors.

```{r Skewed immunizations as factors}
life_new <- life %>% 
  select(-Country, -Year) %>% 
  mutate(Hepatitis_B = ifelse(Hepatitis_B < 86, "<86% Immunized", ">=86% Immunized"),
         Polio = ifelse(Polio < 86, "<86% Immunized", ">=86% Immunized"),
         Diphtheria = ifelse(Diphtheria < 86, "<86% Immunized", ">=86% Immunized"),
         Hepatitis_B = as.factor(Hepatitis_B),
         Polio = as.factor(Polio),
         Diphtheria = as.factor(Diphtheria))

str(life_new)
```



Check the correlation of the numerical variables
```{r Correlation Numeric}
life_numerical <- life_new %>% 
  select_if(is.numeric)

ggcorr(life_numerical, 
       label = T, 
       label_size = 2,
       label_round = 2,
       hjust = 1,
       size = 3, 
       color = "black",
       layout.exp = 5,
       low = "forestgreen", 
       mid = "gray95", 
       high = "darkorange",
       name = "Correlation")
```


Life Expenctancy has a strong positive correlation with Schooling 
and Income_composition_of_resources.

Life_expectancy has a strong negative correlation with Adult_Mortality
which is understandable since if mortality rate in adults is high, Life
exepctancy would be lower.

Population and Measels show almost no correlation with Life_Expectancy.
infant_deaths and under.five_deaths have a correlation of 1 (100%) which would
tell us that there is multicollinearity between them. We should drop one of the
variables, and I believe under 5 deaths should be dropped. 

```{r remove under-five_deaths variable}
life_new<- life_new[-c(10)]
```
## Including Plots

##Check the distribution of the categorical variables


```{r Status Variable, echo=FALSE}
life_new %>% 
  group_by(Status) %>% 
  summarise(count = n()) %>% 
  mutate(percentage = paste0(round(count/sum(count)*100, 2), "%"))



life_plot <-  ggplot(life_new, aes(x=Status, y = Life_expectancy, fill = Status)) +
  geom_boxplot() +
  scale_fill_manual(values=c("forestgreen", "darkorange")) +
  labs(x = "Status", y = "Life Expectancy (Age)") +
  theme(legend.position = "none")


ggplotly(life_plot)

```

According to the boxplots, the distribution of higher life_expectancy resides
in the developed countries, with the plot even showing that the median for developed
countries (2nd Quartile) beind skewed higher towards the 3rd quartile


We want to know if there is any significant difference between the 
average life expectancy in Developed and Developing countries.

```{r ANOVA}
summary(aov(Life_expectancy ~ Status, data = life_new))
```


The ANOVA test tells us that there is a significant difference between the life expectancy
of Developed countries and the life expectancy of Developing countries.


```{r Hepatitis_B Variable, echo=FALSE}
life_new %>% 
  group_by(Hepatitis_B) %>% 
  summarise(count = n()) %>% 
  mutate(percentage = paste0(round(count/sum(count)*100, 2), "%"))




life_plot2 <-  ggplot(life_new, aes(x=Hepatitis_B, y = Life_expectancy, fill = Hepatitis_B)) +
  geom_boxplot() +
  scale_fill_manual(values=c("forestgreen", "darkorange")) +
  labs(x = "Hepatitis B Immunizations", y = "Life Expectancy (Age)") +
  theme(legend.position = "none")


ggplotly(life_plot2)

```

More than 1/3 of countries have less than 86% immunized for Hepatitis B.
The Life Expectancy of the countries with greater than or equal to 86% Immunized is higher than
the countries which have less that 86% Immunized. Note that the median (2nd Quartile) for >=86% immunized
is skewed upwards towards the 3rd quartile.


We want to know if there is any significant difference between the 
average life expectancy in Developed and Developing countries.

```{r ANOVA2}
summary(aov(Life_expectancy ~ Hepatitis_B, data = life_new))
```


The ANOVA test tells us that there is a significant difference between the life expectancy
of countries that immmunized >=86% and the life expectancy of countries that immunized less than 86% for Hep B.



```{r Polio Variable, echo=FALSE}
life_new %>% 
  group_by(Polio) %>% 
  summarise(count = n()) %>% 
  mutate(percentage = paste0(round(count/sum(count)*100, 2), "%"))


life_plot3 <-  ggplot(life_new, aes(x=Polio, y = Life_expectancy, fill = Polio)) +
  geom_boxplot() +
  scale_fill_manual(values=c("forestgreen", "darkorange")) +
  labs(x = "Polio Immunizations", y = "Life Expectancy (Age)") +
  theme(legend.position = "none")


ggplotly(life_plot3)

```

For Polio Immunizations, higher life expectancy resides with the countries 
that have >=86% Immunized.
Note that the median (2nd Quartile) for >=86% immunized
is skewed upwards towards the 3rd quartile.

```{r ANOVA3}
summary(aov(Life_expectancy ~ Polio, data = life_new))
```


According to the ANOVA test, there is a significant difference between the life expectancy
of countries that immmunized >=86% and the life expectancy of countries that immunized less than 86% for Polio.



```{r}

life2 <- life_df_fctr %>%
          drop_na()

life2 %>%
  ggplot() +
  geom_point(
    aes(
      x = log(Population),
      y = Life_expectancy,
      color = WB_region
    )
  ) +
  facet_wrap(~WB_region)


```

There does seem to be an association with repect to the regions having a generally different life expectancy, we may be able to leverage this stat, but I would suppose the region of a country would need to be known.


Here we are looking at the VIFs an we can see there are a number of variables which have a VIF of greater than the 5-10 range, some are in the hundreds. We can remove some of these high multi-colinear values. We can examine the corr plot along with the VIFs to see what we'd like to drop. I will find the things with high correlation and drop identify the one with the higher VIF to drop.

```{r}

life_df_fctr2 <- life2[,c(-1:-3,-24:-32)]  #removing the country name

life_fctr2model <- lm(Life_expectancy~.,data=life_df_fctr2)  # . means all variable not mpg

vif(life_fctr2model)[,3]^2


```





